---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
### Loading Data

First we need to load the data. The data need to be only loaded a single time. That is why we use `cache=TRUE`:

```{r,echo=TRUE}
library(data.table)
zip_file  <- "activity.zip"
text_file <- "activity.csv"
colClasses <- c("numeric","Date","integer") 
df_activity <- read.csv(unz(zip_file, text_file), header = TRUE, sep = ",", colClasses=colClasses,na.strings="NA")
dt_activity <- as.data.table(df_activity)
```

### Number of Steps by day

The next plot shows the histogram of steps. NA values are ignored, and the number of bins was assigned to 20. Notice that the big amount of days with 0 frequency is due in part to the NA values in some days.

```{r, echo=TRUE}
step_daily_average = dt_activity[,list(daily_steps=sum(steps,na.rm = TRUE)),by=date]$daily_steps
hist(
        x=step_daily_average, 
        breaks=20,
        col="blue",
        xlab="Steps",
        ylab = "Frequency",
        main= "Total number of steps by day"     
)
```

Now, we can calculate both the mean and median using:

```{r}
mean_steps   <- mean(step_daily_average)
median_steps <- median(step_daily_average)
```

After running the last code we know that the mean and median are `r mean_steps` and `r median_steps`, respectivelly.

### Average Daily Activity Pattern
The next plot shows the average number of steps by interval, averaged across all days
```{r, echo=TRUE}
dt_temp = dt_activity[,list(average_steps=mean(steps,na.rm = TRUE)),by=interval]
plot(
  x=dt_temp$interval,
  y=dt_temp$average_steps,
  type='l',
  xlab='Interval',
  ylab='Average number of steps by interval'
)
```

Now, for finding the interval with the maximum number of steps averaged across al days we use:
```{r}
max_steps_interval   <- max(dt_temp$average_steps)
max_interval <- dt_temp[average_steps==max_steps_interval,interval]
```

The previous code give us that the 5-minutes interval with the highest number of steps is `r max_interval`

### Missing Values

First we create a new table were we will fill the step missing values using the mean of the 5-minutes interval.

```{r,echo=TRUE}
n_nas <- sum( !is.na( dt_activity$steps ) ) 
dt_activity_fill <- copy(dt_activity)
dt_activity_fill[is.na(steps)]$steps <- with(dt_activity, 
                                             ave(steps, interval, 
                                                 FUN = function(x) mean(x, na.rm = TRUE)))[is.na(dt_activity_fill$steps)]
```

Now, we can repeat the process, for calculating the histogram:

```{r, echo=FALSE}
step_daily_average = dt_activity_fill[,list(daily_steps=sum(steps,na.rm = TRUE)),by=date]$daily_steps
hist(
        x=step_daily_average, 
        breaks=20,
        col="blue",
        xlab="Steps",
        ylab = "Frequency",
        main= "Total number of steps by day"     
)
mean_steps   <- mean(step_daily_average)
median_steps <- median(step_daily_average)
```

We can se that the histogram is totally different, specially for the case of 0 steps. And the reason of this is because there are days
that are not filled at all for the column steps. Finally, the news mean and median are `r mean_steps` and `r median_steps`, respectivelly.

### Weekdays vs Weekends

The next plot will plot the difference between weekends and weekdays

```{r, echo=TRUE}
library(lattice) 
is_weekend_day <- weekdays(dt_activity_fill$date) %in% c("Sunday","Saturday")
dt_activity_fill[, day_type := ifelse(is_weekend_day, "weekend", "weekday")]
dt_temp_fill = dt_activity_fill[,list(average_steps=mean(steps,na.rm = TRUE)),by=list(interval,day_type)]
with(dt_temp_fill, xyplot(average_steps ~ interval | day_type, layout = c(1, 2), type='l',xlab='Interval',ylab="Number of steps"))
```

As we can se there appears to exist some difference at the start of the day. This may due to the fact that on weekends, people normally
wake up later than on weekdays.
